### 2018-08-30 ms/ms_probaweb_for_conversion.tex

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Benjamin Baiser$^{3}$, Jennifer
A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo D. Martinez$^{6}$, Tommi
Nyman$^{7}$, Timothée Poisot$^{2,8}$, Daniel B.
Stouffer$^{9}$, Jason M. Tylianakis$^{9,10}$, Spencer A. Wood$^{11}$, Tomas Roslin$^{12}$ \\

@@

%-----------------
\begin{equation}
	\text{P}(X_{iy},X_{jy},L_{ijy}\lvert E_y)
\end{equation}
%-----------------

@@

%-----------------
\begin{equation}
	P(X_{iy},X_{jy},L_{ijy}\lvert E_y)=P(X_{iy},X_{jy}\lvert E_y)P(L_{ijy}\lvert X_{iy},X_{jy},E_y)
\end{equation}
%-----------------

The first term on the right-hand side of the equation is the probability of
observing the two species co-occurring at location $y$. It corresponds to the
Grinnellian dimension of the niche. The second term represents the probability
that an interaction occurs between species $i$ and $j$, given that they are
co-occurring. This predicate can be refined using information on trait
distribution and trait matching rules . For brevity, we have here avoided
specific references to traits and considered that the interactions are
provided in the data. But one could develop the model further to integrate
such traits, inspired by the framework developed by Gravel et al. 2016. They
proposed to represent the occurence of an interaction between species as a
joint probability event $P(L_{ij}, \mathbf{T}_i, \mathbf{T}_j\lvert E)$, where
$\mathbf{T}_i$ and $\mathbf{T}_j$ are trait vectors for species $i$ and $j$
respectively. In order to develop a trait-based representation of the spatial
variation of interactions, one may thereafter consider how these traits are
related to the environment and how they co-occur. Above, we referred to this
entity as the "metaweb" and note that it corresponds to the Eltonian dimension
of the niche. Below, we will see how this formalism can be directly fit to
empirical data, restricting ourselves to occurrences only and refering to the
above citations for trait-matching and trait-environment relationships. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses concerning spatial variation in network structure (see
the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment,
$P(X_{iy},X_{jy}\lvert E_y)$. In this situation, there are no underlying assumptions
about the ecological processes responsible for co-occurrence. Spatial associations between species could arise
because interactions constrain distribution, where in such case the co-occurrence would be conditional on $L$, or, alternatively, because of
environmental requirements shared between $i$ and $j$ \citep{Pollock2014,
Cazelles2016}. In the former case, species are not independent of each other
and the conditional occurrence must be accounted for explicitly,
$P(X_{iy},X_{jy} \lvert E_y)=P(X_{iy}\lvert E_y,X_{jy})P(X_{jy}\lvert E_y)$. In the latter case,
species are independent, and only the marginal occurrence must be accounted
for, $P( X_{ijy}\lvert E_y)=P(X_{iy} \lvert E_y)P(X_{jy} \lvert E_y)$.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, rather than
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}\lvert X_{iy}=1,X_{jy}=1) = 1$. This approach might be a reasonable
approximation if the spatial or temporal scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity. In this scenario, network variation arises solely from species
distributions.

Second, ecological interactions could also vary with the environment, so that
$P(L_{ijy}\lvert E_y)$. Although it is rare to see a conditional representation of
pairwise ecological interactions, experimental studies have frequently
revealed interactions to be sensitive to the environment. For instance,
\citet{Mckinnon2010} showed that predation risks of shorebirds vary at the
continental scale, decreasing from the south to the north. It is also common
to see increasing top-down control with temperature (e.g. \citealt{Shurin2012,
Gray2016}). Effects of the environment on interactions \citep{Gibert2014} also
propagate up the community and influence network structure
\citep{Tylianakis2007, Woodward2010, Petchey2010a}.

@@

We compared three models for the co-occurrence probability. The first one
directly represents the co-occurrence probability conditional on the local
environment, $P(X_{iy},X_{jy}\lvert E_y)$ (models are listed in Tables 1 and 2).
Hence, this model makes no assumption about the mechanisms driving co-
occurrence for any given environment, and instead uses the information
directly available in the data. It thereby indirectly accounts for the effect
of interactions on co-occurrence, if there are any. The second model considers
independent occurrence of species. In this case, we independently fit
$P(X_{iy} \lvert E_y)$ and $P(X_{jy} \lvert E_y)$, and we then take their product to
derive the probability of co-occurrence. This model should be viewed as a null
hypothesis with respect to the first model, since a comparison between the
respective models will reveal if there is significant spatial association of
the two species beyond a joint response to the shared environment
\citep{Cazelles2016}. Finally, the third model assumes that the probability of
co-occurrence is independent of the environment and thus constant throughout
the landscape. In other words, $P(X_{iy},X_{jy})$ is obtained by simply
counting the number of observed co-occurrences divided by the total number of
observations. Thus, the comparison between the first and third model allows us
to test the hypothesis that co-occurrence is conditional on the environment.
Whenever the environment was included as a covariate in the GLM, we considered
a second-order polynomial term for both temperature and precipitation in
order to account for optima in environmental conditions. There are
consequently five parameters for the first model when fitting a given pair of
species, 10 parameters for the second, and only one for the third model.

Following the same logic, we compared three models of the interaction
probability. The first model conditions the interaction probability on the
local environmental variables, $P(L_{ijy}\lvert X_{iy},X_{jy},E_y)$. Consequently,
the model was fit to the subset of the data where the two species co-occur.
The second model fits the interaction probability independently of the local
environmental variables, $P(L_{ijy}\lvert X_{iy},X_{jy})$. It corresponds to the
number of times the two species were observed to interact when co-occurring,
divided by the number of times that they co-occurred. The third model is an
extreme case evaluated only to test the hypothesis that if two species are
found to interact at least once, then they should interact whenever they co-occur, $P(L_{ijy}\lvert X_{iy},X_{jy})=1$. While not necessarily realistic, this
model tests an assumption commonly invoked in the representation of local
networks from the knowledge of a deterministic metaweb. It represents
potential interactions, rather than realized ones. There are consequently five
parameters for the first model, a single parameter for the second model and no
parameter to evaluate for the third model (where the interaction probability
is fixed by the hypothesis).

We fitted the different models to each pair of species and recorded the
predicted probabilities. The joint probability $P(L_{ijy},X_{iy},X_{jy})$ was
then computed from Eq. 2, and the likelihood of each observation $D_{ijy}$ given the model $\theta$ was computed
as $\mathcal{L}(D_{ijy}\lvert \theta)=P(L_{ij},X_{iy},X_{jy})$ if an
interaction was observed, and as $\mathcal{L}(D_{ijy}\lvert \theta) =
1-P(L_{ijy}, X_{iy}, X_{jy})$ if no interaction was observed. The log-likelihood was summed over the entire dataset to compare the different models
by AIC. We therefore evaluate the likelihood of all local networks, given the
model. Not surprisingly, there was a very large number of species pairs for
which this model could not be computed, as they simply never co-occurred. For
these pairs, we have no information of the interaction probability, and they
were consequently removed from the analysis. The log-likelihood reported
across the entire dataset was therefore summed over all pairs of species observed to co-occur at least once. Interactions between the first (\textit{Salix}) and
second (gallers) trophic layers and those between the second and third
(parasitoids) were considered separately. Finally, we used the full model (in
which both co-occurrence and the interaction are conditional on the
environment) to interpolate species distributions and interaction
probabilities across the entire European continent. We reconstructed the
expected network for each location in a 1 X 1 km grid and computed the
probabilistic connectance following \citet{Poisot2015c}.

@@

We evaluated each model for all pairs of species in order to better understand
the large-scale drivers of network turnover. The results were highly
consistent among trophic levels (\textit{Salix}-gallers and
gallers-parasitoids; Table 3), despite large variability in the fit
of the models to the different pairs of species. This variability was
particularly important for the models of interactions given co-occurrence.
Across all pairs of species, the conditional representation of interactions
performed (Model 2) better than the marginal one (Model 1); that is,
interactions did not occur systematically whenever the two species were found
co-occurring. Hence, in addition to species turnover, the stochastic nature of
interactions contributes to network variability. In total, we recorded 1,173
pairs of interactions, only 290 of which occurred more than five times. Out of
these 290 interactions, 143 were systematically detected whenever the two
species co-occurred. In the instances when species co-occurred, the two
environmental variables considered proved relatively poor predictors of their
interactions (Model 2 vs Model 3). Not surprisingly, for both types of
interactions (\textit{Salix}-galler and galler-parasitoid), the log-likelihood
increased when the environment was considered. However, the extra number of
parameters exceeded the gain in log-likelihood and inflated AIC. Therefore,
the most parsimonious model excluded the effect of the environment. On the
basis of log-likelihood only, co-occurrence was non-independent for both
\textit{Salix}-galler and galler-parasitoid interactions. Thus, according to
AIC, the best model was the one of non-random co-occurrence (Model 3 vs Model
6) for both types of interactions.

@@

Exactly what processes are responsible for the distribution of species
richness at different trophic levels is yet to be established (but see e.g.
\citealt{Roininen2005, Nyman2010, Leppanen2014}), but as a net outcome of
different latitudinal trends across trophic levels, the distribution of co-occurrence and therefore of potential interactions differed between the first
and second layers of feeding links. The correlation between expected
\textit{Salix} and galler richness was 0.73, while it was 0.58 between
gallers and their parasitoids. Therefore, the ratio of herbivores to
\textit{Salix} species is essentially constant across Europe, whereas each
herbivore species is potentially attacked by a richer enemy community higher
latitudes. Consequently, overall connectance peaks in Northern Europe (Fig.
5).

In terms of species interacting with each other, our analysis suggests that
the environment leaves a detectable imprint on species co-occurrence, but only
a slight mark on the occurrence of realized links among species in a specific
place: the probability of finding a given combination of species at a higher
and a lower trophic level at the same site was clearly affected by the
environment, whereas the probability of observing an interaction between the
two was, given co-occurrence, not significantly so. The interactions were  highly
uncertain, as only few pairs of species were systematically interacting when
co-occurring but most not. This applies to the example species
\textit{Phyllocolpa prussica} and \textit{Chrysocharis elongata} (Figs. 3 and
4), but also to all species pairs more generally. For the example species
pair, the full model revealed that the joint probability event (interaction
and co-occurrence) peaks at intermediate temperature and precipitation, simply
because this is where the two species co-occur most often. This does not imply
that species will always interact when they meet – although this is a basic
assumption in most documented metawebs to date (e.g. \citealt{Havens1992,
Wood2015}). Rather, an interaction is better represented as a stochastic
process whose probability is also influenced by the probability with which
species co-occur. What we cannot reliably know is how this stochasticity
splits into different sources of uncertainty – i.e., the extent to which a
species at the higher trophic level runs into a species at the lower level co-occurring at the site, the extent to which this interaction is detected by an
observer collecting a finite sample, or simply the uncertainty arising from
incomplete description of a highly dimensional niche. Future work will be
required to document the relative importance of these sources of uncertainty
in the occurrence of interactions.

@@

Evidence that the structure of ecological networks does vary across habitats
(e.g. \citealt{Tylianakis2007}), over environmental gradients
\citep{Lurgi2012} and in time \citep{Simanonok2014} is accumulating
rapidly. It is not clear, however, to what extent the turnover of network
structure is driven by a systematic change in species composition or of
pairwise interactions \citep{Poisot2012, Poisot2015a}. Our model comparison of
host-parasite interactions revealed that most of the turnover is driven by
species-specific responses to the environment, impacting species richness, and
that co-occurrence was mostly independent. Further, the occurrence of interactions
among host and parasite is highly stochastic even when both are present, and
not predictable by the variables considered by us. We know that interactions
vary with the environment in other systems, for instance, herbivory
\citep{Shurin2012, Baskett2018} and predation \citep{Mckinnon2010,
Legagneux2014} are often found to increase with temperature, resulting in
spatial variation of trophic cascades \citep{Gray2016}. What remains unclear,
however, is the extent to which such variation is driven by a turnover of
species composition along gradients, or a turnover of the interactions. Here
we found that interactions vary substantially but non-predictably along the
annual temperature and the precipitation gradient. The lack of detectable
signal may be due to our choice of covariates. Indeed, a previous study focusing on a subset of the system studied here identified habitat characteristics as the primary drivers of
interactions \citep{Nyman2015}. New investigations with other systems will
thus be required to challenge this result. Under all circumstances,
documenting the relationship between the environment and the occurrence of
interactions at continental scales is critical for understanding how large-
scale variation of trophic regulation influences community dynamics and
ecosystem functioning \citep{Harfoot2014}.

The framework we provide complements methods to compute network beta-diversity
presented by \citealt{Poisot2012} and \citealt{Poisot2018}. The total network
turnover is partitioned into interaction turnover and species turnover – which
in our approach would correspond to the three terms of Eq. 2, respectively.
Importantly, one could easily derive the different $\beta$ of
\citealt{Poisot2012} using the fitted probabilities. For instance, for a pair
of species $i$ and $j$ and sites $x$ and $y$, the expectation for the fraction
$a$ for the interactions would be
$P(L_{ijx}\lvert X_{ix},X_{ijx})P(L_{ijx}\lvert X_{ix},X_{ijx})$. The same could be
computed for fractions $b$, and therefore the expected beta-diversity for an
entire network could be recomputed this way. The novel feature of the approach
presented here is then that the different contributions to network turnover
can be attributed to different hypotheses (as performed here), in order to
evaluate for instance the role of environmental heterogeneity or species co-
distribution on network turnover.

We restricted our analysis to the effect of co-occurrence on ecological
interactions, neglecting the inverse of the problem. We did not investigate in
depth the drivers of co-occurrence and simply took it for granted from the
data. Co-occurrence was indeed different from the expectation of independent
species distributions. It thus begs the question of whether, once
environmental effects on species-specific distribution have been accounted
for, interactions come with significant effects on co-occurrence? We could
rephrase this problem by asking whether the fundamental niche differs from the
realized niche, and how this applies to our framework. For example, we have
considered above simply the co-occurrence probability, $P(X_{iy},X_{jy}\lvert E_y)$,
which could be expanded as $P(X_{iy}\lvert X_{jy},E_y) P(X_{jy}\lvert  E_y )$. After some
re-arrangement of Eq. 2, the marginal occurrence probability, $P(X_{jy}\lvert E_y
)$, could be considered as a species distribution model taking into account
the interaction between these species (in the same spirit as done by Gonzalez-
Salazar et al. 2013). This derivation would however critically depend on a
strong \emph{a priori} expectation of the conditional probability of observing
a species given the distribution of the other species. This assumption seems
reasonable for some situations, such as a parasitoid species that requires a
host to develop. On the other hand, \citet{Cazelles2016b} found that the
strength of this association is often rather weak if not inexistant (for
instance, with the example pair presented in Table 1). The lack of an
association could simply arise when the parasitoid is generalist enough that
it is not constrained to track the distribution of any single/given host
\citep{Cazelles2015}.

The approach we presented could easily be expanded to account for traits, in
order to derive a complete and more mechanistic representation of the niche.
In this context, both the Eltonian and Grinnellian dimensions of the niche
could be represented. As mentioned earlier, \citealt{Gravel2016} proposed to
represent the occurrence of interactions based on some trait-matching rules
(given by a function of the type $P(L_{ij})\lvert \mathbf{T}_i, \mathbf{T}_j, E)$)
and the co-distribution of traits (corresponding to $P(\mathbf{T}_i,
\mathbf{T}_j\lvert E)$). The key  insight from the formalism proposed by
\citealt{Gravel2016} is that the co-distribution of traits will impact
variation in network structure. As a result, it was found that network
properties often co-vary with functional structure and functional diversity
\citealp{Laigle2018}. The next step in the development of this framework will
be to develop the models for the co-distribution of traits, for instance
following Ovaskainen et al. 2017. This may facilitate the accessibility and
the applicability of the framework, since the basic unit of analysis is no
longer the species, but rather a combination of traits, thereby facilitating
the statistical analysis by sharing information among species.

@@

%\begin{landscape}
\begin{table}[]
\centering

\caption{Example of data used to perform the analysis. Traditional species distribution models represent the occurrence of single species as a function of the environment, with models such as $X_i­~Temp$ performed one species after the other. Some also tried to represent both the abiotic and biotic niche, with models such as $X_i­~Temp + X_j$ (e.g. \citealt{GonzalezSalasar2013}), while the new joint species distribution models \citep{Ovaskainen2017} consider the entire community at once ($\mathbf{X}~Temp$) and aims at representing co-distribution ($X_{ij}$). Here the focus is on the occurrence of interactions, $L_{ij}$. The full model (see model 3, Tables 1 and 2) considers the result of two submodels, one representing the occurrence of interactions given co-occurrence, $L_{ij}~Temp$ (evaluated only on the subset of data where $X_{ij}=1$), and the other representing the co-occurrence, $X_{ij}~Temp$.}

\begin{tabular}{llllllll}
\hline

>>>>

\textbf{Authors:} Dominique Gravel (1,2,*), Benjamin Baiser (3), Jennifer
A. Dunne (4), Jens-Peter Kopelke (5), Neo D. Martinez (6), Tommi
Nyman (7), Timothée Poisot (2,8), Daniel B.
Stouffer (9), Jason M. Tylianakis (9,10), Spencer A. Wood (11), Tomas Roslin (12) \\

@@

%-----------------
\begin{equation}
	\text{P}(X_{iy},X_{jy},L_{ijy}\mid E_y)
\end{equation}
%-----------------

@@

%-----------------
\begin{equation}
	P(X_{iy},X_{jy},L_{ijy}\mid E_y)=P(X_{iy},X_{jy}\mid E_y)P(L_{ijy}\mid X_{iy},X_{jy},E_y)
\end{equation}
%-----------------

The first term on the right-hand side of the equation is the probability of
observing the two species co-occurring at location $y$. It corresponds to the
Grinnellian dimension of the niche. The second term represents the probability
that an interaction occurs between species $i$ and $j$, given that they are
co-occurring. This predicate can be refined using information on trait
distribution and trait matching rules . For brevity, we have here avoided
specific references to traits and considered that the interactions are
provided in the data. But one could develop the model further to integrate
such traits, inspired by the framework developed by Gravel et al. 2016. They
proposed to represent the occurence of an interaction between species as a
joint probability event $P(L_{ij}, \mathbf{T}_i, \mathbf{T}_j\mid E)$, where
$\mathbf{T}_i$ and $\mathbf{T}_j$ are trait vectors for species $i$ and $j$
respectively. In order to develop a trait-based representation of the spatial
variation of interactions, one may thereafter consider how these traits are
related to the environment and how they co-occur. Above, we referred to this
entity as the "metaweb" and note that it corresponds to the Eltonian dimension
of the niche. Below, we will see how this formalism can be directly fit to
empirical data, restricting ourselves to occurrences only and refering to the
above citations for trait-matching and trait-environment relationships. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses concerning spatial variation in network structure (see
the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment,
$P(X_{iy},X_{jy}\mid E_y)$. In this situation, there are no underlying assumptions
about the ecological processes responsible for co-occurrence. Spatial associations between species could arise
because interactions constrain distribution, where in such case the co-occurrence would be conditional on $L$, or, alternatively, because of
environmental requirements shared between $i$ and $j$ \citep{Pollock2014,
Cazelles2016}. In the former case, species are not independent of each other
and the conditional occurrence must be accounted for explicitly,
$P(X_{iy},X_{jy} \mid E_y)=P(X_{iy}\mid E_y,X_{jy})P(X_{jy}\mid E_y)$. In the latter case,
species are independent, and only the marginal occurrence must be accounted
for, $P( X_{ijy}\mid E_y)=P(X_{iy} \mid E_y)P(X_{jy} \mid E_y)$.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, rather than
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}\mid X_{iy}=1,X_{jy}=1) = 1$. This approach might be a reasonable
approximation if the spatial or temporal scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity. In this scenario, network variation arises solely from species
distributions.

Second, ecological interactions could also vary with the environment, so that
$P(L_{ijy}\mid E_y)$. Although it is rare to see a conditional representation of
pairwise ecological interactions, experimental studies have frequently
revealed interactions to be sensitive to the environment. For instance,
\citet{Mckinnon2010} showed that predation risks of shorebirds vary at the
continental scale, decreasing from the south to the north. It is also common
to see increasing top-down control with temperature (e.g. \citealt{Shurin2012,
Gray2016}). Effects of the environment on interactions \citep{Gibert2014} also
propagate up the community and influence network structure
\citep{Tylianakis2007, Woodward2010, Petchey2010a}.

@@

We compared three models for the co-occurrence probability. The first one
directly represents the co-occurrence probability conditional on the local
environment, $P(X_{iy},X_{jy}\mid E_y)$ (models are listed in Tables 1 and 2).
Hence, this model makes no assumption about the mechanisms driving co-
occurrence for any given environment, and instead uses the information
directly available in the data. It thereby indirectly accounts for the effect
of interactions on co-occurrence, if there are any. The second model considers
independent occurrence of species. In this case, we independently fit
$P(X_{iy} \mid E_y)$ and $P(X_{jy} \mid E_y)$, and we then take their product to
derive the probability of co-occurrence. This model should be viewed as a null
hypothesis with respect to the first model, since a comparison between the
respective models will reveal if there is significant spatial association of
the two species beyond a joint response to the shared environment
\citep{Cazelles2016}. Finally, the third model assumes that the probability of
co-occurrence is independent of the environment and thus constant throughout
the landscape. In other words, $P(X_{iy},X_{jy})$ is obtained by simply
counting the number of observed co-occurrences divided by the total number of
observations. Thus, the comparison between the first and third model allows us
to test the hypothesis that co-occurrence is conditional on the environment.
Whenever the environment was included as a covariate in the GLM, we considered
a second-order polynomial term for both temperature and precipitation in order
to account for optima in environmental conditions. There are consequently five
parameters for the first model when fitting a given pair of species, 10
parameters for the second, and only one for the third model.

Following the same logic, we compared three models of the interaction
probability. The first model conditions the interaction probability on the
local environmental variables, $P(L_{ijy}\mid X_{iy},X_{jy},E_y)$. Consequently,
the model was fit to the subset of the data where the two species co-occur.
The second model fits the interaction probability independently of the local
environmental variables, $P(L_{ijy}\mid X_{iy},X_{jy})$. It corresponds to the
number of times the two species were observed to interact when co-occurring,
divided by the number of times that they co-occurred. The third model is an
extreme case evaluated only to test the hypothesis that if two species are
found to interact at least once, then they should interact whenever they co-
occur, $P(L_{ijy}\mid X_{iy},X_{jy})=1$. While not necessarily realistic, this
model tests an assumption commonly invoked in the representation of local
networks from the knowledge of a deterministic metaweb. It represents
potential interactions, rather than realized ones. There are consequently five
parameters for the first model, a single parameter for the second model and no
parameter to evaluate for the third model (where the interaction probability
is fixed by the hypothesis).

We fitted the different models to each pair of species and recorded the
predicted probabilities. The joint probability $P(L_{ijy},X_{iy},X_{jy})$ was
then computed from Eq. 2, and the likelihood of each observation $D_{ijy}$
given the model $\theta$ was computed as
$\mathcal{L}(D_{ijy}\mid \theta)=P(L_{ij},X_{iy},X_{jy})$ if an interaction was
observed, and as $\mathcal{L}(D_{ijy}\mid \theta) = 1-P(L_{ijy}, X_{iy}, X_{jy})$
if no interaction was observed. The log-likelihood was summed over the entire
dataset to compare the different models by AIC. We therefore evaluate the
likelihood of all local networks, given the model. Not surprisingly, there was
a very large number of species pairs for which this model could not be
computed, as they simply never co-occurred. For these pairs, we have no
information of the interaction probability, and they were consequently removed
from the analysis. The log-likelihood reported across the entire dataset was
therefore summed over all pairs of species observed to co-occur at least once.
Interactions between the first (\textit{Salix}) and second (gallers) trophic
layers and those between the second and third (parasitoids) were considered
separately. Finally, we used the full model (in which both co-occurrence and
the interaction are conditional on the environment) to interpolate species
distributions and interaction probabilities across the entire European
continent. We reconstructed the expected network for each location in a 1 X 1
km grid and computed the probabilistic connectance following
\citet{Poisot2015c}.

@@

We evaluated each model for all pairs of species in order to better understand
the large-scale drivers of network turnover. The results were highly
consistent among trophic levels (\textit{Salix}-gallers and gallers-
parasitoids; Table 3), despite large variability in the fit of the models to
the different pairs of species. This variability was particularly important
for the models of interactions given co-occurrence. Across all pairs of
species, the conditional representation of interactions performed (Model 2)
better than the marginal one (Model 1); that is, interactions did not occur
systematically whenever the two species were found co-occurring. Hence, in
addition to species turnover, the stochastic nature of interactions
contributes to network variability. In total, we recorded 1,173 pairs of
interactions, only 290 of which occurred more than five times. Out of these
290 interactions, 143 were systematically detected whenever the two species
co-occurred. In the instances when species co-occurred, the two environmental
variables considered proved relatively poor predictors of their interactions
(Model 2 vs Model 3). Not surprisingly, for both types of interactions
(\textit{Salix}-galler and galler-parasitoid), the log-likelihood increased
when the environment was considered. However, the extra number of parameters
exceeded the gain in log-likelihood and inflated AIC. Therefore, the most
parsimonious model excluded the effect of the environment. On the basis of
log-likelihood only, co-occurrence was non-independent for both
\textit{Salix}-galler and galler-parasitoid interactions. Thus, according to
AIC, the best model was the one of non-random co-occurrence (Model 3 vs Model
6) for both types of interactions.

@@

Exactly what processes are responsible for the distribution of species
richness at different trophic levels is yet to be established (but see e.g.
\citealt{Roininen2005, Nyman2010, Leppanen2014}), but as a net outcome of
different latitudinal trends across trophic levels, the distribution of co-
occurrence and therefore of potential interactions differed between the first
and second layers of feeding links. The correlation between expected
\textit{Salix} and galler richness was 0.73, while it was 0.58 between gallers
and their parasitoids. Therefore, the ratio of herbivores to \textit{Salix}
species is essentially constant across Europe, whereas each herbivore species
is potentially attacked by a richer enemy community higher latitudes.
Consequently, overall connectance peaks in Northern Europe (Fig. 5).

In terms of species interacting with each other, our analysis suggests that
the environment leaves a detectable imprint on species co-occurrence, but only
a slight mark on the occurrence of realized links among species in a specific
place: the probability of finding a given combination of species at a higher
and a lower trophic level at the same site was clearly affected by the
environment, whereas the probability of observing an interaction between the
two was, given co-occurrence, not significantly so. The interactions were
highly uncertain, as only few pairs of species were systematically interacting
when co-occurring but most not. This applies to the example species
\textit{Phyllocolpa prussica} and \textit{Chrysocharis elongata} (Figs. 3 and
4), but also to all species pairs more generally. For the example species
pair, the full model revealed that the joint probability event (interaction
and co-occurrence) peaks at intermediate temperature and precipitation, simply
because this is where the two species co-occur most often. This does not imply
that species will always interact when they meet – although this is a basic
assumption in most documented metawebs to date (e.g. \citealt{Havens1992,
Wood2015}). Rather, an interaction is better represented as a stochastic
process whose probability is also influenced by the probability with which
species co-occur. What we cannot reliably know is how this stochasticity
splits into different sources of uncertainty – i.e., the extent to which a
species at the higher trophic level runs into a species at the lower level co-
occurring at the site, the extent to which this interaction is detected by an
observer collecting a finite sample, or simply the uncertainty arising from
incomplete description of a highly dimensional niche. Future work will be
required to document the relative importance of these sources of uncertainty
in the occurrence of interactions.

@@

Evidence that the structure of ecological networks does vary across habitats
(e.g. \citealt{Tylianakis2007}), over environmental gradients
\citep{Lurgi2012} and in time \citep{Simanonok2014} is accumulating rapidly.
It is not clear, however, to what extent the turnover of network structure is
driven by a systematic change in species composition or of pairwise
interactions \citep{Poisot2012, Poisot2015a}. Our model comparison of host-
parasite interactions revealed that most of the turnover is driven by species-
specific responses to the environment, impacting species richness, and that
co-occurrence was mostly independent. Further, the occurrence of interactions
among host and parasite is highly stochastic even when both are present, and
not predictable by the variables considered by us. We know that interactions
vary with the environment in other systems, for instance, herbivory
\citep{Shurin2012, Baskett2018} and predation \citep{Mckinnon2010,
Legagneux2014} are often found to increase with temperature, resulting in
spatial variation of trophic cascades \citep{Gray2016}. What remains unclear,
however, is the extent to which such variation is driven by a turnover of
species composition along gradients, or a turnover of the interactions. Here
we found that interactions vary substantially but non-predictably along the
annual temperature and the precipitation gradient. The lack of detectable
signal may be due to our choice of covariates. Indeed, a previous study
focusing on a subset of the system studied here identified habitat
characteristics as the primary drivers of interactions \citep{Nyman2015}. New
investigations with other systems will thus be required to challenge this
result. Under all circumstances, documenting the relationship between the
environment and the occurrence of interactions at continental scales is
critical for understanding how large- scale variation of trophic regulation
influences community dynamics and ecosystem functioning \citep{Harfoot2014}.

The framework we provide complements methods to compute network beta-diversity
presented by \citealt{Poisot2012} and \citealt{Poisot2018}. The total network
turnover is partitioned into interaction turnover and species turnover – which
in our approach would correspond to the three terms of Eq. 2, respectively.
Importantly, one could easily derive the different $\beta$ of
\citealt{Poisot2012} using the fitted probabilities. For instance, for a pair
of species $i$ and $j$ and sites $x$ and $y$, the expectation for the fraction
$a$ for the interactions would be
$P(L_{ijx}\mid X_{ix},X_{ijx})P(L_{ijx}\mid X_{ix},X_{ijx})$. The same could be
computed for fractions $b$, and therefore the expected beta-diversity for an
entire network could be recomputed this way. The novel feature of the approach
presented here is then that the different contributions to network turnover
can be attributed to different hypotheses (as performed here), in order to
evaluate for instance the role of environmental heterogeneity or species co-
distribution on network turnover.

We restricted our analysis to the effect of co-occurrence on ecological
interactions, neglecting the inverse of the problem. We did not investigate in
depth the drivers of co-occurrence and simply took it for granted from the
data. Co-occurrence was indeed different from the expectation of independent
species distributions. It thus begs the question of whether, once
environmental effects on species-specific distribution have been accounted
for, interactions come with significant effects on co-occurrence? We could
rephrase this problem by asking whether the fundamental niche differs from the
realized niche, and how this applies to our framework. For example, we have
considered above simply the co-occurrence probability, $P(X_{iy},X_{jy}\mid E_y)$,
which could be expanded as $P(X_{iy}\mid X_{jy},E_y) P(X_{jy}\mid  E_y )$. After some
re-arrangement of Eq. 2, the marginal occurrence probability, $P(X_{jy}\mid E_y
)$, could be considered as a species distribution model taking into account
the interaction between these species (in the same spirit as done by Gonzalez-
Salazar et al. 2013). This derivation would however critically depend on a
strong \emph{a priori} expectation of the conditional probability of observing
a species given the distribution of the other species. This assumption seems
reasonable for some situations, such as a parasitoid species that requires a
host to develop. On the other hand, \citet{Cazelles2016b} found that the
strength of this association is often rather weak if not inexistant (for
instance, with the example pair presented in Table 1). The lack of an
association could simply arise when the parasitoid is generalist enough that
it is not constrained to track the distribution of any single/given host
\citep{Cazelles2015}.

The approach we presented could easily be expanded to account for traits, in
order to derive a complete and more mechanistic representation of the niche.
In this context, both the Eltonian and Grinnellian dimensions of the niche
could be represented. As mentioned earlier, \citealt{Gravel2016} proposed to
represent the occurrence of interactions based on some trait-matching rules
(given by a function of the type $P(L_{ij})\mid \mathbf{T}_i, \mathbf{T}_j, E)$)
and the co-distribution of traits (corresponding to $P(\mathbf{T}_i,
\mathbf{T}_j\mid E)$). The key  insight from the formalism proposed by
\citealt{Gravel2016} is that the co-distribution of traits will impact
variation in network structure. As a result, it was found that network
properties often co-vary with functional structure and functional diversity
\citealp{Laigle2018}. The next step in the development of this framework will
be to develop the models for the co-distribution of traits, for instance
following Ovaskainen et al. 2017. This may facilitate the accessibility and
the applicability of the framework, since the basic unit of analysis is no
longer the species, but rather a combination of traits, thereby facilitating
the statistical analysis by sharing information among species.

@@

% %\begin{landscape}
% \begin{table}[]
% \centering

% \caption{Example of data used to perform the analysis. Traditional species distribution models represent the occurrence of single species as a function of the environment, with models such as $X_i­~Temp$ performed one species after the other. Some also tried to represent both the abiotic and biotic niche, with models such as $X_i­~Temp + X_j$ (e.g. \citealt{GonzalezSalasar2013}), while the new joint species distribution models \citep{Ovaskainen2017} consider the entire community at once ($\mathbf{X}~Temp$) and aims at representing co-distribution ($X_{ij}$). Here the focus is on the occurrence of interactions, $L_{ij}$. The full model (see model 3, Tables 1 and 2) considers the result of two submodels, one representing the occurrence of interactions given co-occurrence, $L_{ij}~Temp$ (evaluated only on the subset of data where $X_{ij}=1$), and the other representing the co-occurrence, $X_{ij}~Temp$.}

% \begin{tabular}{llllllll}
% \hline

% Location & $Sp_i$ & $Sp_j$ & $X_i$ & $X_j$ & $X_{ij}$ & $L_{ij}$ & Av. temp \\ \hline
% Kühkopf 		& Phyllocolpa prussica  & Chrysocharis elongata & 1 & 1 & 1 & 1 &  8.25 \\
% Kühkopf 		& Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 1 & 0 & 0 &  8.25 \\
% Ötztal 			& Phyllocolpa prussica  & Chrysocharis elongata & 0 & 0 & 0 & 0 & -1.50 \\
% Ötztal 			& Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 0 & 0 & 0 & -1.50 \\
% Kilpisjärvi		& Phyllocolpa prussica  & Chrysocharis elongata & 0 & 1 & 0 & 0 & -1.92 \\
% Kilpisjärvi		& Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 1 & 0 & 0 & -1.92 \\
% Wüstensachsen	& Phyllocolpa prussica  & Chrysocharis elongata & 1 & 1 & 1 & 0 &  5.58 \\
% Wüstensachsen	& Phyllocolpa leucapsis & Chrysocharis elongata & 1 & 1 & 1 & 1 &  5.58 \\
% Steinamoen      & Phyllocolpa prussica  & Chrysocharis elongata & 0 & 1 & 0 & 0 &  3.17 \\
% Steinamoen      & Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 1 & 0 & 0 &  3.17 \\

% \hline
% \end{tabular}
% \end{table}
% %\end{landscape}
